---
title: "Analyzing a Superstore dataset"
author: "Laia Porcar, Luis Marcos López, Philippe Robert"
date: "7/9/2022"
output:
  html_document:
    toc: yes
    df_print: kable
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```

## Installs and libraries

```{r, results="hide", message = FALSE}
# Installs
#install.packages("caret")
#install.packages("car")
#install.packages("ggplot2")
#install.packages("devtools")
#install.packages("ggcorrplot")
#devtools::install_github("laresbernardo/lares")
#devtools::install_github("ProcessMiner/nlcor")
#install.packages("ISLR2")
#devtools::install_github("kassambara/ggpubr")
#install.packages("zoo")
#install.packages("ROSE")
#install.packages("e1071")
#install.packages("klaR")
#install.packages("treemapify")
#install.packages("glmnet")

# Libraries
library(knitr)
library(tidyverse)
library(ggplot2)
library(forcats)
library(dplyr)
library(ggcorrplot)
library(lares)
library(treemapify)
library(MASS)
library(devtools) 
library(ISLR2)
library(ggpubr)
library(zoo)
library(reshape2)
library(ROSE) # for undersampling/oversampling
library(e1071) # for naive Bayes
library(caret)
library(klaR)
library(car) # for calculation of VIF in multiple linear regression
library(glmnet)
library(choroplethr)
library(choroplethrMaps)
library(kableExtra)
library(magrittr)

```

## Importing Dataset

```{r results = 'asis'}
# Read the dataset
df <- read.csv("~/Users/34669/Documents/Data Science/Statistical Learning/Sample-Superstore.csv",header=TRUE)
#kable(df[1:5,])

df[1:15,] %>%
  kable(format = "html", col.names = colnames(df)) %>%
  column_spec(c(2:21), width_min = "0.8in")  %>%
  kable_styling() %>%
  kableExtra::scroll_box(width = "100%", height = "300px")


```

## Data cleaning

### Basic data cleaning: info, nan, removing/adding columns, duplicates...

Summary of the data:

```{r}
summary(df)
str(df)
```

Check if there are Na and NAN:

```{r}
is.null(df)
sum(is.na(df))
sum(is.nan(as.matrix(df)))
```

Unique values in each column:

```{r}
cat(paste("Sub.Category: ", length(unique(df[["Sub.Category"]])) ), sep="\n")
cat(paste("Category: ", length(unique(df[["Category"]]))), sep="\n")
cat(paste("Region: ", length(unique(df[["Region"]]))), sep="\n")
cat(paste("State: ", length(unique(df[["State"]]))), sep="\n")
cat(paste("City: ", length(unique(df[["City"]]))), sep="\n")
cat(paste("Ship.Mode: ", length(unique(df[["Ship.Mode"]]))), sep="\n")
cat(paste("Customer.ID: ", length(unique(df[["Customer.ID"]]))), sep="\n")
```

Number of duplicates:

```{r}
sum(duplicated(df))
```

Removing useless columns, i.e. Customer name, Country and Postal Code:

```{r}
df = df[-c(1,7,9,12)]
```

Convert date columns to date format:

```{r}
df$Order.Date <- as.Date(df$Order.Date, format = "%m/%d/%Y")
df$Ship.Date <- as.Date(df$Ship.Date, format = "%m/%d/%Y")
```

Create useful columns:

```{r}
df$Processing.Time <- df$Ship.Date - df$Order.Date
df$Processing.Time <- as.numeric(df$Processing.Time)
df$Gross.Margin <- (df$Profit/df$Sales)*100
df$Order.Year <- format(df$Order.Date,"%Y")
df$Order.Year <- as.numeric(df$Order.Year)                                
df$Order.Month <- format(df$Order.Date,"%m")
df$Order.Month <- as.numeric(df$Order.Month)
df$Order.Year_month <- format(df$Order.Date,"%Y-%m")
df$Order.Year_month <- as.Date(paste(df$Order.Year_month,"-15",sep=""))
df$Order.Date <- as.numeric(df$Order.Date)
```

Let's see again the final structure of the data after adding new columns and removing useless columns

```{r}
str(df)
```

### Outliers detection

Let's have a look again at the summary of columns that could potentially have some outliers.

```{r}
summary(df$Sales)
summary(df$Quantity)
summary(df$Discount)
summary(df$Profit)
summary(df$Processing.Time)
summary(df$Order.Year)
summary(df$Gross.Margin)
summary(df$Order.Date)
```

Let's plot the numerical variable to have a quick look at their distribution.

```{r warning = FALSE, message = FALSE, fig.height=32, fig.width=32}
hst_Sales <- ggplot(df) + aes(x = Sales) + 
geom_histogram(color="white", fill="#80b1d3", bins = 100) + 
labs(title="", x="Sales", y="Frequency") +
theme_minimal()+ 
theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), title = element_text(size = 16))

hst_Quantity <- ggplot(df) + aes(x = Quantity) + 
geom_histogram(color="white", fill="#80b1d3", bins = 14) + 
labs(title="", x="Quantity", y="Frequency") + 
theme_minimal()+ 
theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
  title = element_text(size = 16))

hst_Discount <- ggplot(df) + aes(x = Discount) + 
geom_histogram(color="white", fill="#80b1d3", bins = 9) + 
labs(title="", x="Discount", y="Frequency") + 
theme_minimal()+ 
theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
  title = element_text(size = 16))

hst_Profit <- ggplot(df) + aes(x = Profit) + 
geom_histogram(color="white", fill="#80b1d3", bins = 120) + 
labs(title="", x="Profit", y="Frequency") + 
theme_minimal()+ 
theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
  title = element_text(size = 16))

hst_Processing <- ggplot(df) + aes(x = Processing.Time) + 
geom_histogram(color="white", fill="#80b1d3", bins = 8)  + 
labs(title="", x="Processing time", y="Frequency") +
theme_minimal()+  
theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
  title = element_text(size = 16))

hst_Year <- ggplot(df) + aes(x = Order.Year) + 
geom_histogram(color="white", fill="#80b1d3", bins = 4)  + 
labs(title="", x="Years", y="Frequency") + 
theme_minimal()+ 
theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
  title = element_text(size = 16))

################################################################################

require(gridExtra)
grid.arrange(hst_Sales, hst_Quantity, hst_Year, hst_Discount, hst_Processing, hst_Profit, ncol = 2, nrow = 3)
```

Let's look at the profit histogram more in detail. First separating positive and negative values and doing a log transformation.

```{r message = FALSE, fig.height=26, fig.width=32}
df_positive_profit <- df %>% filter(Profit >= 0)
df_negative_profit <- df %>% filter(Profit < 0)

plot1 <- ggplot(df_positive_profit) + aes(x = log(Profit)) + 
  geom_histogram(color="white", fill="#80b1d3",  bins = 12) + 
  labs(title="", x="Log(Pos. Profit)", y="Frequency") + 
  theme_minimal()+ 
  theme(axis.text=element_text(size=24), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28), title = element_text(size = 16))

plot2 <- ggplot(df_negative_profit) + aes(x = log(-Profit)) + 
  geom_histogram(color="white", fill="#80b1d3", bins = 12) + 
  labs(title="", x="Log(Neg. Profit)", y="Frequency") + 
  theme_minimal()+ 
  theme(axis.text=element_text(size=24), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28), title = element_text(size = 16))

# Also the sales histogram with the log transformation:
plot3 <- ggplot(df) + aes(x = log(Sales)) + 
  geom_histogram(color="white", fill="#80b1d3", bins = 30) + 
  labs(title="", x="log(Sales)", y="Frequency") + 
  theme_minimal()+ 
  theme(axis.text=element_text(size=24), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28), title = element_text(size = 16))


# Now removing the central high values, so we can see better the extremes of the histogram.
df_profit <- df[!(df$Profit < 600 & df$Profit >= -600 ),]

plot4 <- ggplot(df_profit) + aes(x = Profit) + 
  geom_histogram(color="white", fill="#80b1d3") + 
  labs(title="", x="Profit (omitting central values)", y="Frequency") + 
  theme_minimal()+ 
  theme(axis.text=element_text(size=24), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28), title = element_text(size = 16))

################################################################################


require(gridExtra)
grid.arrange(plot1, plot2, plot3, plot4, ncol = 2, nrow = 2)
```

Boxplots

```{r message = FALSE, fig.height=26, fig.width=32}
plot1 <- ggplot(df, aes(y = Sales)) + 
  geom_boxplot(color="black", fill="#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), title = element_text(size = 16))

plot2 <- ggplot(df, aes(y = Quantity)) + 
  geom_boxplot(color="black", fill="#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), title = element_text(size = 16))

plot3 <- ggplot(df, aes(y = Discount)) + 
  geom_boxplot(color="black", fill="#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), title = element_text(size = 16))

plot4 <- ggplot(df, aes(y = Profit)) + 
  geom_boxplot(color="black", fill="#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), title = element_text(size = 16))

plot5 <- ggplot(df, aes(y = Processing.Time)) + 
  geom_boxplot(color="black", fill="#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), title = element_text(size = 16))

# Alternatively we can also separate the quantitative variable per the qualitative one, for example:

plot6 <- ggplot(df, aes(x = Category, y = Discount)) + 
  geom_boxplot(color="black", fill="#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), title = element_text(size = 16))

plot7 <- ggplot(df, aes(x = Region, y = Sales)) + 
  geom_boxplot(color="black", fill="#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), title = element_text(size = 16))

################################################################################

require(gridExtra)
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, ncol = 4, nrow = 2)
```

## Exploratory Data Analysis

```{r}
str(df)
```

### Analysis of discrete variables

```{r, message = FALSE, fig.height=22, fig.width=32}
plotdata <- df %>% count(Sub.Category) %>% arrange(desc(Sub.Category)) %>%
      mutate(prop = round(n*100/sum(n), 1), lab.ypos = cumsum(prop) - 0.5*prop)

plotdata$label <- paste0(plotdata$Sub.Category, "\n", round(plotdata$prop), "%")

plot1 <- ggplot(plotdata, aes(x = "", y = prop, fill = Sub.Category)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 1.4, direction = -1) +
  theme(legend.position = "FALSE", axis.text=element_text(size=32))+
  scale_y_continuous(breaks=cumsum(plotdata$prop) - plotdata$prop / 2, 
  labels= plotdata$label) + labs(y = "", x = "", title = "") + theme_minimal()+   theme(axis.text=element_text(size=24), axis.title.x = element_text(size =28),   axis.title.y = element_text(size = 28), legend.text = element_text(size =24),
  legend.title = element_text(size =28))


################################################################################

df_cat <- df %>% group_by(Category, Sub.Category) %>% summarize(count=n())

plot2 <- ggplot(df_cat, aes(area=count, label=Sub.Category, fill=Category, subgroup=Category))+
    geom_treemap()+
    geom_treemap_subgroup_border(colour = "white", size = 5)+
    geom_treemap_subgroup_text(color='white', alpha=0.3, size=48,
    place='center', fontface='bold')+
    geom_treemap_text(color='white', place='center', size=34) +
    theme(legend.text = element_text(size = 24), 
    legend.title = element_text(size = 28))

################################################################################

require(gridExtra)
grid.arrange(plot1, plot2, ncol = 2, nrow = 1)
```

```{r, message = FALSE, fig.height=36, fig.width=32}
top_states <- tail(names(sort(table(df$State))),15)
df_top_states <- df %>% filter(df$State %in% top_states)
plot1 <- ggplot(df_top_states, aes(x = State, fill = Category)) + 
  geom_bar(position = "stack", bins = 15) + 
  labs(title="", x="State", y="Orders") + 
  theme_minimal()+ 
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45),
  axis.title.x = element_text(size =28),   axis.title.y=element_text(size=28),
  plot.title = element_text(size = 24, hjust = 0.5), 
  legend.text = element_text(size = 24), legend.title = element_text(size =28)) +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

top_cities <- tail(names(sort(table(df$City))),15)
df_top_cities <- df %>% filter(df$City %in% top_cities)
plot2 <- ggplot(df_top_cities, aes(x = City, fill = Category)) + 
  geom_bar(position = "stack", bins = 15) + 
  labs(title="", x = "City", y="Orders") + 
  theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size=28),
    plot.title = element_text(size = 24, hjust = 0.5), 
    legend.text = element_text(size = 24), legend.title = element_text(size=28))+
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))


################################################################################

plot3 <- ggplot(data = df, aes(x = Sub.Category, y = stat(count), fill = Segment, 
    label = scales::percent(prop.table(stat(count))))) +
  geom_bar(position = "stack", bins = 3) + 
  labs(title="", x = "Sub-category", y="Orders") + 
  theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size=28),
    axis.text.x = element_text(angle=45), axis.title.y = element_text(size=28),
    plot.title = element_text(size = 24, hjust = 0.5),  
    legend.text = element_text(size = 24), legend.title = element_text(size=28))+
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

df2 <- df %>% group_by(Category)
plot4 <- ggplot(data = df2,aes(Category, fill=`Ship.Mode`))+
  geom_bar(position = 'stack')+
  labs(title='', x='Category', y='Orders') +
  theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28),
    plot.title = element_text(size = 24, hjust = 0.5),  
    legend.text = element_text(size = 24), legend.title = element_text(size =28)) +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

df2 <- df %>% group_by(Segment)
plot5 <- ggplot(data = df2, aes(Segment, fill=`Ship.Mode`))+
  geom_bar(position = 'stack')+
  labs(title='', x='Segment', y='Orders') +
  theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28),  
    plot.title = element_text(size = 24, hjust = 0.5),  
    legend.text = element_text(size = 24), legend.title = element_text(size=28))+
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

df2 <-df %>% group_by(Region)
plot6 <- ggplot(data = df2, aes(Region, fill=`Ship.Mode`))+
  geom_bar(position = 'stack')+
  theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 24, hjust = 0.5),  
    legend.text = element_text(size = 24), legend.title = element_text(size=28))+
  labs(title='', x='Region', y='Orders') +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

# Purchases, sales and benefits for years and months. In what months do people usually buy more? 
# Are the sales generally increasing over time? Are there more sales on Christmas, Black Friday...?

plot7 <- ggplot(df, aes(x=Order.Month, fill = Category))+ 
  geom_bar(position = "stack", bins = 12)+ 
  labs(title="", x = "Month", y="Orders")+ 
  theme_minimal() +  
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 24, hjust = 0.5),  
    legend.text = element_text(size = 24), legend.title = element_text(size=28))+
  scale_x_discrete(limits= seq(1,12))+ 
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

# Is there any preference on shipment methods over the years?

plot8 <- ggplot(df, aes(x=Order.Year, fill = Ship.Mode))+ 
  geom_bar(position = "stack",  bins = 4)+ 
  labs(title="", x = "Year", y="Orders")+ 
  theme_minimal()+ 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 24, hjust = 0.5),  
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

require(gridExtra)
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, ncol = 2, nrow = 4)
```

```{r, message = FALSE, fig.height=28, fig.width=32}
plot1 <- ggplot(df, aes(x=Order.Month, y= Profit, fill = Category))+
  geom_bar(stat = "identity") + labs(title="", x="Month", y="Profit") + 
  theme_minimal() + theme(axis.text=element_text(size=24), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
    plot.title = element_text(size = 16, hjust = 0.5), 
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) + 
  scale_x_discrete(limits= seq(1,12)) + scale_y_continuous(label = scales::dollar) +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

plot2 <- ggplot(df, aes(x=Order.Month, y=Sales, fill = Category))+ 
  geom_bar(stat = "identity", bins = 12) + labs(title="", x="Month", y="Sales") + 
  theme_minimal() + theme(axis.text=element_text(size=24), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
    plot.title = element_text(size = 16, hjust = 0.5), 
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) + 
  scale_x_discrete(limits= seq(1,12)) + scale_y_continuous(label = scales::dollar) +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

plot3 <- ggplot(df, aes(x=Segment, y=Sales, fill = Region)) + 
  geom_bar(stat = "identity") + labs(x="Segment", y="Sales") + theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))+
  scale_y_continuous(label = scales::dollar)
  
################################################################################

plot4 <- ggplot(df, aes(x=Region, y=Sales, fill = Category)) + 
  geom_bar(stat = "identity") + labs(x="Region", y="Sales") + theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))+
  scale_y_continuous(label = scales::dollar)

################################################################################

require(gridExtra)
grid.arrange(plot2, plot1, plot3, plot4, ncol = 2, nrow = 2)
```

```{r, message = FALSE, fig.height=22, fig.width=32}
df_subcat <- df %>% group_by(Sub.Category) %>% summarise(prof = sum(Profit))
df_subcat <- df_subcat[order(df_subcat$prof),]
top_subcat_neg <- head(df_subcat$Sub.Category, 5)
top_subcat_pos <- tail(df_subcat$Sub.Category, 5)
df_top_pos <- df %>% filter(Sub.Category %in% top_subcat_pos)

plot_top_pos <- ggplot(df_top_pos, aes(x = Sub.Category, y = Profit, fill = Segment)) + 
  geom_bar(stat="identity", position = "stack", bins = 5) + 
  labs(x = "Subcategory", y= "Profit") + 
  theme_minimal() + scale_y_continuous(label = scales::dollar) +
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 38, hjust = 0.5), 
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  ggtitle("Top subcategories with positive profit") +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))
  
  
################################################################################

df_top_neg <- df %>% filter(Sub.Category %in% top_subcat_neg)

plot_top_neg <- ggplot(df_top_neg, aes(x = Sub.Category, y = Profit, fill = Segment)) + 
  geom_bar(stat = "identity", position = "stack", bins = 5) + 
  labs(x = "Subcategory", y= "Profit") + scale_y_continuous(label = scales::dollar) +
  theme_minimal() +
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  ggtitle("Top subcategories with negative profit") +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))

################################################################################

require(gridExtra)
grid.arrange(plot_top_pos, plot_top_neg, ncol = 2, nrow = 1)
```

Total profit by state

```{r, message = FALSE, fig.height=8, fig.width=8}
df_mapusa <- df %>% group_by(State) %>% summarize(profitss = sum(Profit))
df_mapusa$region <- tolower(df_mapusa$State)
df_mapusa$value <- df_mapusa$profitss
continental_us_states <- names(table(df_mapusa$region))

state_choropleth(df_mapusa, num_colors=9, zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") +
  labs(fill = "Total profit") 

```

Total orders by state

```{r, message = FALSE, fig.height=8, fig.width=8}
df_mapusa <- df %>% group_by(State)
df_mapusa <- df_mapusa %>% summarize(orders = n())
df_mapusa$region <- tolower(df_mapusa$State)
df_mapusa$value <- df_mapusa$orders
continental_us_states <- names(table(df_mapusa$region))

state_choropleth(df_mapusa, num_colors=9, zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") +
  labs(fill = "Total orders") 
```

Total clients by state (keep in mind often some clients buy in different states)

```{r, message = FALSE, fig.height=8, fig.width=8}
df_2 <- df %>% group_by(Customer.ID, State) %>% summarize(prueba = n())
df_mapusa <- df_2 %>% group_by(State) %>% summarize(clients = n())
df_mapusa$region <- tolower(df_mapusa$State)
df_mapusa$value <- df_mapusa$clients
continental_us_states <- names(table(df_mapusa$region))

state_choropleth(df_mapusa, num_colors=9, zoom = continental_us_states) +
 scale_fill_brewer(palette="YlOrBr") +
 labs(fill = "N. Clients") 
```

```{r, message = FALSE, fig.height=18, fig.width=32}
Sales_per_year <- df %>%  group_by(Order.Year) %>% summarise(s = sum(Sales))

Profit_per_year <- df %>% group_by(Order.Year) %>% summarise(p = sum(Profit))

Sales <- Sales_per_year$s
Profit <- Profit_per_year$p
Years <- c("2014", "2015", "2016", "2017")

df_growthsales_profit <- data.frame(Sales, Profit, Years)
df_growthsales_profit2 <- melt(df_growthsales_profit, id.vars="Years")

plot_year <- ggplot(df_growthsales_profit2, aes(x=Years, y=value, fill=variable))+
  geom_bar(stat="identity", position="dodge") + 
  labs(title="Sales and Profit per year", x = "Years", y="USD") + 
  theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) + 
  scale_fill_manual(values = c("#da6474", "#80b1d3", "#fdb462", "#7fc97f"))


###############################################################################

Order_preholiday <- df %>% filter(Order.Month %in% 1:10) %>% 
                    group_by(Order.Year) %>% summarise(ord = n())

Order_holiday <- df %>% filter(Order.Month %in% 11:12) %>% 
                 group_by(Order.Year) %>% summarise(ord1= n())  

Pre_holiday <- Order_preholiday$ord
Holiday <- Order_holiday$ord1
Years <- c("2014", "2015", "2016", "2017")

df_growthorder <- data.frame(Pre_holiday, Holiday, Years)
df_growthorder2 <- melt(df_growthorder, id.vars='Years')

plot_hol <- ggplot(df_growthorder2, aes(x=Years, y=value, fill=variable)) +
    geom_bar(stat='identity', position='dodge') + 
    labs(title="Orders Pre-Holiday vs Holiday", x = "Years", y="Orders") + 
    theme_minimal() +
    theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
      axis.title.y = element_text(size = 28), 
      plot.title = element_text(size = 38, hjust = 0.5),
      legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
    scale_fill_manual(values = c("#da6474", "#80b1d3", "#fdb462", "#7fc97f" ))

################################################################################

require(gridExtra)
grid.arrange(plot_year, plot_hol, ncol = 2, nrow = 1)
```

```{r, message = FALSE, fig.height=14, fig.width=32}
Technology_sales <- df %>% filter(str_detect(Category, 'Technology')) %>%
  group_by(Order.Year) %>% summarise(Sales_technology = sum(Sales))

 Technology_profit <- df %>% filter(str_detect(Category, "Technology")) %>%
   group_by(Order.Year) %>% summarise(Profit_technology = sum(Profit))

Years <- c("2014", "2015", "2016", "2017")
Sales <- Technology_sales$Sales_technology
Profit <- Technology_profit$Profit_technology

df_technology_sp <- data.frame(Sales, Profit, Years)
df_technology_sp2 <- melt(df_technology_sp, id.vars="Years")

plot_technology <- ggplot(df_technology_sp2, aes(x=Years, y=value, fill=variable)) +
  geom_bar(stat="identity", position="dodge") + 
  labs(title="Sales and Profit for Technology", x = "Years", y="USD") + 
  theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  scale_fill_manual(values = c("#da6474", "#80b1d3", "#fdb462", "#7fc97f" ))

############################################################################

Furniture_sales <- df %>% filter(str_detect(Category, 'Furniture')) %>%
  group_by(Order.Year) %>% summarise(Sales_furniture = sum(Sales))

Furniture_profit <- df %>% filter(str_detect(Category, "Furniture")) %>%
  group_by(Order.Year) %>% summarise(Profit_furniture = sum(Profit))

Profit <- Furniture_profit$Profit_furniture
Sales <- Furniture_sales$Sales_furniture

df_furniture_sp <- data.frame(Sales, Profit, Years)
df_furniture_sp2 <- melt(df_furniture_sp, id.vars="Years")

plot_furniture <- ggplot(df_furniture_sp2, aes(x=Years, y=value, fill=variable)) +
  geom_bar(stat="identity", position="dodge") + 
  labs(title="Sales and Profit for Furniture", x = "Years", y="USD") + 
  theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  scale_fill_manual(values = c("#da6474", "#80b1d3", "#fdb462", "#7fc97f" ))

#############################################################################

Office_supplies_sales <- df %>% filter(str_detect(Category, 'Office Supplies')) %>%
  group_by(Order.Year) %>% summarise(Sales_office_supplies = sum(Sales))

Office_supplies_profit <- df %>% filter(str_detect(Category, "Office Supplies")) %>%
   group_by(Order.Year) %>% summarise(Profit_office_supplies = sum(Profit))

Profit <- Office_supplies_profit$Profit_office_supplies
Sales <- Office_supplies_sales$Sales_office_supplies

df_office_supplies_sp <- data.frame(Sales, Profit, Years)
df_office_supplies_sp2 <- melt(df_office_supplies_sp, id.vars="Years")

plot_office <- ggplot(df_office_supplies_sp2, aes(x=Years, y=value, fill=variable)) +
   geom_bar(stat="identity", position="dodge") + 
   labs(title="Sales and Profit for Office Supplies", x = "Years", y="USD") +
   theme_minimal() +
   theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
   scale_fill_manual(values = c("#da6474", "#80b1d3", "#fdb462", "#7fc97f" ))

################################################################################

require(gridExtra)
grid.arrange(plot_technology, plot_furniture, plot_office, ncol = 3, nrow = 1)
```

```{r, message = FALSE, fig.height=28, fig.width=32}
df1 <- df %>% group_by(Segment,Category) %>% summarise(n=median(Profit))
plot1 <- ggplot(df1, aes(Segment, Category, fill=n))+
  scale_fill_distiller( direction = 1)+ geom_tile(color='white')+
  geom_text(aes(label=paste(round(n,0),'$')), color = 'black', size=10)+
  labs(title='', fill='Median Profit') + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28))

################################################################################

df2 <- df %>% group_by(Region,Category) %>% summarise(n=median(Profit))
plot2 <- ggplot(df2, aes(Region, Category, fill=n))+
  scale_fill_distiller( direction = 1)+ geom_tile(color='white')+
  geom_text(aes(label=paste(round(n,0),'$')), color = 'black', size=10)+
  labs(title='', fill='Median Profit') + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28))

################################################################################

# df3 <- df %>% mutate(Discount = cut_width(Discount,0.2,boundary=0))
df3 <- df %>% mutate(Discount = cut(Discount, c(0, 0.1, 0.2, 0.8), include.lowest = TRUE))
df3 <- df3 %>% group_by(Discount, Sub.Category) %>% summarise(n = mean(Quantity))

plot3 <- ggplot(df3, aes(Discount, Sub.Category, fill= n))+
  scale_fill_distiller( direction = 1) + geom_tile(color='white')+
  geom_text(aes(label=paste(round(n,2))), color = 'black', size=10)+
  labs(title='', x = "Discount", fill='Avg. Quantity') + theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28))


################################################################################
df4 <- df %>% mutate(Discount = cut_width(Discount,0.15,boundary=0))
df4 <- df4 %>% group_by(Discount)

plot4 <- ggplot(df4, aes(x=Discount, y= Profit, fill = Category))+
  geom_bar(stat = "identity", position = "stack") + labs(title="", x="Discount", y="Profit") + 
  theme_minimal()+
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  scale_y_continuous(label = scales::dollar) +
  scale_fill_manual(values = c("#da6474", "#80b1d3", "#fdb462", "#7fc97f" ))

################################################################################

require(gridExtra)
grid.arrange(plot1, plot2, plot3, plot4, ncol = 2, nrow = 2)
```

```{r, message = FALSE, fig.height=16, fig.width=32}
df_preholiday <- df %>% filter(Order.Month %in% 1:10) %>% group_by(Order.Year_month) %>%
  summarise(m1 = mean(Sales), mx1 = max(Sales), mn = min(Sales) ,su = sum(Sales))

plot1 <- ggplot(df_preholiday, aes(x = Order.Year_month, y = su)) +
  geom_point(color="#80b1d3", size = 6, alpha=.9) +   theme_minimal()+
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28))+
  scale_y_continuous(label = scales::dollar) +
  labs(x = "Time", y = "Total sales per month from January to October")+
  geom_smooth(method = "lm") + stat_cor(method = "pearson", size = 10)

################################################################################

df_holiday <- df %>% filter(Order.Month %in% 11:12) %>% group_by(Order.Year_month) %>%
  summarise(m1 = mean(Sales), mx1 = max(Sales), mn1 = min(Sales) ,su1 = sum(Sales))

plot2 <- ggplot(df_holiday, aes(x = Order.Year_month, y = su1)) +
  geom_point(color="#80b1d3", size = 6, alpha=.9) + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28)) +
  scale_y_continuous(label = scales::dollar) +
  labs(x = "Time", y = "Total sales per month for November and December")+
  geom_smooth(method = "lm") + stat_cor(method = "pearson", size = 10)

###############################################################################

require(gridExtra)
grid.arrange(plot1, plot2, ncol = 2, nrow = 1)
```

```{r, message = FALSE, fig.height=32, fig.width=32}
by_yearmonth <- df %>% group_by(Order.Year_month)
sales_year_tot <- by_yearmonth %>% summarize(Sales_year_tot=sum(Sales))

plot1 <- ggplot(sales_year_tot, aes(x = Order.Year_month, y = Sales_year_tot)) +
  geom_line(size = 1.5, color = "lightgrey") + geom_smooth() + 
  geom_point(size = 6, color = "#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5)) +
  geom_point(size = 5, color = "#80b1d3") + labs(y = "Total sales", x = "Time")+
  scale_y_continuous(label = scales::dollar)

################################################################################


df_fullyear <- df %>% group_by(Order.Year_month) %>%
  summarise(m2 = mean(Sales), mx2 = max(Sales), mn2 = min(Sales) ,su2 = sum(Sales))

df_fullyear$Moving_average <- rollapply(df_fullyear$su2, 3, mean, align = "right", fill = NA)

df_fullyear[1,"Moving_average"] = df_fullyear[1,"su2"]
df_fullyear[2,"Moving_average"] = sum(df_fullyear[1,"su2"], df_fullyear[2,"su2"])/2 
#since the moving average is with the past 3 months, the first two row are NA.
# the above lines find a workaround

plot2 <- ggplot(df_fullyear, aes(x = Order.Year_month, y = Moving_average)) +
  geom_line(size = 1.5, color = "lightgrey") + geom_smooth() +
  geom_point(size = 6, color = "#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5)) +
  labs(y = "Avg. sales of last 3 months", x = "Time") + 
  scale_y_continuous(label = scales::dollar)

################################################################################

by_yearmonth <- df %>% group_by(Order.Year_month)
profit_year_tot <- by_yearmonth %>% summarize(Profit_year_tot=sum(Profit))

plot3 <- ggplot(profit_year_tot, aes(x = Order.Year_month, y = Profit_year_tot)) +
  geom_line(size = 1.5, color = "lightgrey") + geom_smooth() +
  geom_point(size = 6, color = "#80b1d3") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5)) +
  labs(y = "Total profit", x = "Time")+
  scale_y_continuous(label = scales::dollar)

################################################################################


df_fullyear <- df %>% group_by(Order.Year_month) %>%
  summarise(m2 = mean(Profit), mx2 = max(Profit), mn2 = min(Profit) ,su2 = sum(Profit))

df_fullyear$Moving_average <- rollapply(df_fullyear$su2, 3, mean, align = "right", fill = NA)

df_fullyear[1,"Moving_average"] = df_fullyear[1,"su2"]
df_fullyear[2,"Moving_average"] = sum(df_fullyear[1,"su2"], df_fullyear[2,"su2"])/2 
#since the moving average is with the past 3 months, the first two row are NA.
# the above lines find a workaround

plot4 <- ggplot(df_fullyear, aes(x = Order.Year_month, y = Moving_average)) +
  geom_line(size = 1.5, color = "lightgrey") + geom_smooth() +
  geom_point(size = 6, color = "#80b1d3") + theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
    axis.title.y = element_text(size = 28), 
    plot.title = element_text(size = 16, hjust = 0.5)) +
  labs(y = "Avg. profit of last 3 months", x = "Time")+
  scale_y_continuous(label = scales::dollar)

################################################################################

require(gridExtra)
grid.arrange(plot1, plot2, plot3, plot4, ncol = 2, nrow = 2)
```

### Top 10 clients analysis

```{r, message = FALSE, fig.height=32, fig.width=32}
top_buyers <- tail(names(sort(table(df$Customer.ID))), 10)
df_top_buyers <- df %>% filter(df$Customer.ID %in% top_buyers)

segment_purch <- df_top_buyers %>% group_by(Customer.ID, Segment) %>% 
  summarize(Sales = sum(Sales), Profit = sum(Profit), Number.Orders = n())
#kable(segment_purch[1:10,])
segment_purch[1:10,] %>%
  kable() %>%
  kable_styling(full_width = F)

plot_top <- ggplot(df_top_buyers, aes(x = Customer.ID, fill = Category)) + 
  geom_bar(position = "stack",bins = 10) + 
  labs(x = "Clients", y="Orders") +
  theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),     plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) + 
  ggtitle("Top 10 clients according to purchases") +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f")) 
  

################################################################################

plot_top_profit <- ggplot(df_top_buyers, aes(x = Customer.ID, y = Profit, fill = Category)) + 
  geom_bar(stat="identity", position = "stack", bins = 10) + 
  labs(x = "Clients", y="Profit") + 
  theme_minimal() + 
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  ggtitle("Top 10 clients according to purchases") +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))+
  scale_y_continuous(label = scales::dollar)

################################################################################

df_profit <- df %>% group_by(Customer.ID) %>% summarise(prof = sum(Profit))
df_profit <- df_profit[order(df_profit$prof),]
top_profit_neg <- head(df_profit$Customer.ID, 10)
top_profit_pos <- tail(df_profit$Customer.ID, 10)

df_top_pos <- df %>% filter(Customer.ID %in% top_profit_pos)

segment_prof_pos <- df_top_pos %>% group_by(Customer.ID, Segment) %>% 
  summarize(Sales = sum(Sales), Profit = sum(Profit), Number.Orders = n())
segment_prof_pos[1:10,] %>%
  kable() %>%
  kable_styling(full_width = F)


plot_top_pos <- ggplot(df_top_pos, aes(x = Customer.ID, y = Profit, fill = Category)) + 
  geom_bar(stat="identity", position = "stack", bins = 10) + 
  labs(x = "Clients", y="Profit") + 
  theme_minimal() +
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  ggtitle("Top 10 clients with positive profit") +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))+
  scale_y_continuous(label = scales::dollar)

################################################################################

df_top_neg <- df %>% filter(df$Customer.ID %in% top_profit_neg)

segment_prof_neg <- df_top_neg %>% group_by(Customer.ID, Segment) %>% 
  summarize(Sales = sum(Sales), Profit = sum(Profit), Number.Orders = n())
segment_prof_neg[1:10,] %>%
  kable() %>%
  kable_styling(full_width = F)


plot_top_neg <- ggplot(df_top_neg, aes(x = Customer.ID, y = Profit, fill = Category)) + 
  geom_bar(stat = "identity", position = "stack", bins = 10) + 
  labs(x = "Clients", y="Profit") + 
  theme_minimal() +
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  ggtitle("Top 10 clients with negative profit") +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))+
  scale_y_continuous(label = scales::dollar)

################################################################################

df_sales <- df %>% group_by(Customer.ID) %>% summarise(sales = sum(Sales))
df_sales <- df_profit[order(df_sales$sales),]
top_sales <- tail(df_sales$Customer.ID, 10)

df_top_sales <- df %>% filter(Customer.ID %in% top_sales)

segment_sales <- df_top_sales %>% group_by(Customer.ID, Segment) %>% 
  summarize(Sales = sum(Sales), Profit = sum(Profit), Number.Orders = n())
segment_sales[1:10,] %>%
  kable() %>%
  kable_styling(full_width = F)


plot_sales <- ggplot(df_top_sales, aes(x = Customer.ID, y = Sales, fill = Category)) + 
  geom_bar(stat="identity", position = "stack", bins = 10) + 
  labs(x = "Clients", y="Sales") + 
  theme_minimal() +
  theme(axis.text=element_text(size=24), axis.text.x = element_text(angle=45), 
    axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
    plot.title = element_text(size = 38, hjust = 0.5),
    legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  ggtitle("Top 10 clients according to sales") +
  scale_fill_manual(values = c("#da6474","#80b1d3", "#fdb462", "#7fc97f"))+
  scale_y_continuous(label = scales::dollar)

################################################################################

require(gridExtra)
grid.arrange(plot_top, plot_top_profit, plot_top_pos, plot_top_neg, plot_sales, ncol = 2, nrow = 3)
```

### Analysis of continuous variables

#### Profit vs Sales

```{r, message = FALSE, fig.height=5, fig.width=7}
ggplot(df, aes(x = Sales,  y = Profit, color=Category)) +
  geom_point(size = 2, alpha=.8) +
  scale_x_continuous(label = scales::dollar) +
  scale_y_continuous(label = scales::dollar) +
  labs(x = "Sales", y = "Profit")+
  theme_minimal() + theme(axis.text=element_text(size=6), 
  axis.title.x = element_text(size = 10), axis.title.y = element_text(size = 10),
  legend.text = element_text(size = 6), legend.title = element_text(size=10))
```

```{r, message = FALSE, fig.height=32, fig.width=32}
# Cut width of Discount
df_disc <- df %>% mutate(Discount=cut_width(Discount,0.10,boundary=0))

df1 <- df_disc 
plot1 <- ggplot(df1, aes(x = Sales,  y = Profit, color=Discount)) +
  geom_point(size = 5, alpha=.8) +
  facet_wrap(~Sub.Category, scales = "free") +
  scale_x_continuous(label = scales::dollar) +
  scale_y_continuous(label = scales::dollar) +
  labs(x = "Sales", y = "Profit") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
  axis.text.x = element_text(angle=45),
  axis.title.y = element_text(size = 28), strip.text = element_text(size = 28),
  legend.text = element_text(size = 24), legend.title = element_text(size=28))

require(gridExtra)
grid.arrange(plot1, ncol = 1, nrow = 1)
```

#### Profit vs Time

```{r, message = FALSE, fig.height=16, fig.width=32}
# Profit over time
df_plot <- df %>% group_by(Order.Year_month)
df_plot <- df_plot %>% summarize(profitt = sum(Profit))

plot1 <- ggplot(df_plot, aes(x = Order.Year_month, y = profitt)) +
  geom_point(color="#80b1d3", size = 6, alpha=.8) +
  scale_y_continuous(label = scales::dollar) +
  labs(x = "Month", y = "Total profit")+
  geom_smooth(method = "lm") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), plot.title = element_text(size = 38, 
  hjust = 0.5), legend.text = element_text(size = 24), 
  legend.title = element_text(size=28)) + stat_cor(method = "pearson", size = 10)

######################################################################

# Orders over time
df_plot <- df %>% group_by(Order.Year_month)
df_plot <- df_plot %>% summarize(orders = n())

plot2 <- ggplot(df_plot, aes(x = Order.Year_month, y = orders)) +
  geom_point(color="#80b1d3", size = 6, alpha=.8) +
  scale_y_continuous(label = scales::dollar) +
  labs(x = "Month", y = "Total orders")+
  geom_smooth(method = "lm") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), plot.title = element_text(size = 38, 
  hjust = 0.5), legend.text = element_text(size = 24), 
  legend.title = element_text(size=28)) + stat_cor(method = "pearson", size = 10)

#######################################################################

require(gridExtra)
grid.arrange(plot1, plot2, ncol = 2, nrow = 1)
```

```{r, message = FALSE, fig.height=32, fig.width=32}
#BY SUBCATEGORIES
df_plot_SC <- df %>% group_by(Sub.Category, Order.Year_month)
df_plot_SC <- df_plot_SC %>% summarize(profit = sum(Profit))

ggplot(df_plot_SC, aes(x = Order.Year_month,  y = profit)) +
  geom_point(color="#80b1d3", size = 5, alpha=.8) +
  facet_wrap(~Sub.Category, scales = "free") +
  scale_y_continuous(label = scales::dollar) +
  #geom_hline(yintercept=0, color="#da6474", size=0.5)+
  labs(x = "Time", y = "Profit")+
  geom_smooth(method = "lm")+
  theme_minimal()+ theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
  strip.text = element_text(size = 28), 
  legend.text = element_text(size = 24), legend.title = element_text(size=28)) +
  stat_cor(method = "pearson", size = 7)

#BY STATE
df_plot_SC <- df %>% group_by(State, Order.Year_month)
df_plot_SC <- df_plot_SC %>% summarize(profit = sum(Profit))

ggplot(df_plot_SC, aes(x = Order.Year_month,  y = profit)) +
  geom_point(color="#80b1d3", size = 5, alpha=.8) +
  facet_wrap(~State, scales = "free") +
  scale_y_continuous(label = scales::dollar) +
  #geom_hline(yintercept=0, color="#da6474", size=0.5)+
  labs(x = "Time", y = "Profit")+
  geom_smooth(method = "lm")+
  theme_minimal() + theme(axis.text=element_text(size=24), 
  axis.title.x = element_text(size = 28), axis.title.y = element_text(size = 28),
  strip.text = element_text(size = 28), axis.text.x = element_text(angle=45),
  legend.text = element_text(size = 24),  legend.title = element_text(size=28))
```

#### Profit vs Month

```{r, message = FALSE, fig.height=32, fig.width=32}
#BY SUBCATEGORIES
df_plot_SC <- df %>% group_by(Sub.Category, Order.Month)
df_plot_SC <- df_plot_SC %>% summarize(profit = sum(Profit))

ggplot(df_plot_SC, aes(x = Order.Month,  y = profit)) +
  geom_point(color="#80b1d3", size = 5, alpha=.8) +
  facet_wrap(~Sub.Category, scale = "free") +
  scale_y_continuous(label = scales::dollar) +
  #geom_hline(yintercept=0, color="#da6474", size=0.5)+
  labs(x = "Month", y = "Profit") +
  geom_smooth(method = "lm") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), strip.text = element_text(size = 28))+
  stat_cor(method = "pearson", size = 7)

#BY STATE
df_plot_SC <- df %>% group_by(State, Order.Month)
df_plot_SC <- df_plot_SC %>% summarize(profit = sum(Profit))

ggplot(df_plot_SC, aes(x = Order.Month,  y = profit)) +
  geom_point(color="#80b1d3", size = 5, alpha=.8) +
  facet_wrap(~State, scale = "free") +
  scale_y_continuous(label = scales::dollar) +
  #geom_hline(yintercept=0, color="#da6474", size=0.5)+
  labs(x = "Month", y = "Profit") +
  geom_smooth(method = "lm") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), strip.text = element_text(size = 28))
```

#### Profit vs Discount

```{r, message = FALSE, fig.height=32, fig.width=32}
#BY SUBCATEGORIES
df_plot_SC <- df %>% group_by(Sub.Category, Discount)
df_plot_SC <- df_plot_SC %>% summarize(profit = sum(Profit))

ggplot(df_plot_SC, aes(x = Discount,  y = profit)) +
  geom_point(color="#80b1d3", size = 5, alpha=.8) +
  facet_wrap(~Sub.Category, scale = "free") +
  scale_y_continuous(label = scales::dollar) +
  #geom_hline(yintercept=0, color="#da6474", size=0.5)+
  labs(x = "Discount", y = "Profit")+
  geom_smooth(method = "lm") +  theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28),
  axis.title.y = element_text(size = 28), strip.text = element_text(size = 28)) +
  stat_cor(method = "pearson", size = 7)
```

#### Profit vs Quantity

```{r, message = FALSE, fig.height=32, fig.width=32}
#BY SUBCATEGORIES
df_plot_SC <- df %>% group_by(Sub.Category, Quantity)
df_plot_SC <- df_plot_SC %>% summarize(profit = sum(Profit))

ggplot(df_plot_SC, aes(x = Quantity,  y = profit)) +
  geom_point(color="#80b1d3", size = 5, alpha=.8) +
  facet_wrap(~Sub.Category, scale = "free") +
  scale_y_continuous(label = scales::dollar) +
  #geom_hline(yintercept=0, color="#da6474", size=0.5)+
  labs(x = "Quantity", y = "Profit") +
  geom_smooth(method = "lm") + theme_minimal() + 
    theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), strip.text = element_text(size = 28)) +  stat_cor(method = "pearson", size = 7)
```

#### Quantity vs Discount

```{r, message = FALSE, fig.height=22, fig.width=32}
#BY SUBCATEGORIES
df_plot_SC <- df %>% group_by(Sub.Category, Discount)
df_plot_SC <- df_plot_SC %>% summarize(quantity = mean(Quantity))

ggplot(df_plot_SC, aes(x = Discount,  y = quantity)) +
  geom_point(color="#80b1d3", size = 5, alpha=.8) +
  facet_wrap(~Sub.Category, scale = "free") +
  #geom_hline(yintercept=0, color="#da6474", size=0.5)+
  labs(x = "Discount", y = "Quantity") +
  geom_smooth(method = "lm") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), strip.text = element_text(size = 28)) +
  stat_cor(method = "pearson", size = 7)


```

```{r, message = FALSE, fig.height=14, fig.width=32}

 #BY SEGMENT
df_plot_SC <- df %>% group_by(Segment, Discount)
df_plot_SC <- df_plot_SC %>% summarize(quantity = mean(Quantity))

ggplot(df_plot_SC, aes(x = Discount,  y = quantity)) +
  geom_point(color="#80b1d3", size = 6, alpha=.8) +
  facet_wrap(~Segment, scale = "free") +
  #geom_hline(yintercept=0, color="#da6474", size=0.5)+
  labs(x = "Discount", y = "Quantity") +
  geom_smooth(method = "lm") + theme_minimal() +
  theme(axis.text=element_text(size=24), axis.title.x = element_text(size = 28), 
  axis.title.y = element_text(size = 28), strip.text = element_text(size = 28)) + stat_cor(method = "pearson", size = 7)

```

### Correlation analysis

```{r, message = FALSE, fig.height=5, fig.width=5}
dat <- df[, c("Sales","Profit","Gross.Margin" , "Discount", "Quantity", "Processing.Time", "Order.Year", "Order.Month")]
colnames(dat) <- c("Sales", "Profit","Gross.Margin", "Disc.", "Quantity", "Processing.Time", "Year", "Month")

corr_cross(dat, max_pvalue = 0.05, top = 10, color=c("#da6474", "#80b1d3"))

```

```{r, message = FALSE, fig.height=5, fig.width=5}
dat <- df[,c("Sales","Profit","Gross.Margin" , "Discount", "Quantity", "Processing.Time", "Order.Year", "Order.Month")]
colnames(dat) <- c("Sales", "Profit","Gross.Margin", "Disc.", "Quantity", "Processing.Time", "Year", "Month")

df_corr <- dplyr::select_if(dat, is.numeric)
corr_coef <- cor(df_corr, use="complete.obs")
round(corr_coef,2)

ggcorrplot(corr_coef, hc.order = TRUE, type = "lower", lab = TRUE)

```

## Linear Regression

### Profit vs Sales for Envelopes

```{r, message = FALSE}
df_envelop <- df %>% filter(`Sub.Category`=='Envelopes')

## Divide training ans test dataset
row.number <- sample(1:nrow(df_envelop), 0.8*nrow(df_envelop))
train = df_envelop[row.number,]
test = df_envelop[-row.number,]
dim(train)
dim(test)
```

```{r, message = FALSE, fig.height=5, fig.width=5}
lm.fit <- lm(Profit ~ Sales, data = train)
summary(lm.fit)
coef(lm.fit)
confint(lm.fit)

ggplot(train, aes(x = Sales, y = Profit)) +
  geom_point(color="#80b1d3", size = 3, alpha=.9) +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) +
  labs(x = "Sales", y = "Profit") + theme_minimal() +
  theme(axis.text=element_text(size=14), axis.title.x = element_text(size = 18), 
  axis.title.y = element_text(size = 18)) + geom_abline(slope = coef(lm.fit)[["Sales"]], 
  intercept = coef(lm.fit)[["(Intercept)"]])

```

```{r, message = FALSE}
pred <- predict(lm.fit, newdata = test)
rmse <- sqrt(sum((pred - test$Profit)^2)/length(test$Profit))
normalized_rmse <- 100* rmse/(max(train$Profit)-min(train$Profit))
cat(paste("RMSE =", rmse))
cat(paste("\nNormalized RMSE =", normalized_rmse, "%"))
```

RMSE explains on an average how much of the predicted value will be from the actual value. Based on RMSE = 3.278, we can conclude that on an average predicted value will be off by 3.278 from the actual value.

Diagnostic plots

```{r, message = FALSE, fig.height=10, fig.width=10}
par(mfrow = c(2, 2))
plot(lm.fit)
```

```{r, message = FALSE, fig.height=10, fig.width=10}
par(mfrow = c(2, 2))
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

### Gross Margin vs Discount

```{r, message = FALSE}
df_disc <- df %>% group_by(Discount) %>% summarize(margin = mean(Gross.Margin))

## Divide training ans test dataset
row.number <- sample(1:nrow(df_disc), 0.8*nrow(df_disc))
train = df_disc[row.number,]
test = df_disc[-row.number,]
dim(train)
dim(test)
```

```{r, message = FALSE, fig.height=5, fig.width=5}
lm.fit <- lm(margin ~ Discount, data = train)
summary(lm.fit)
coef(lm.fit)
confint(lm.fit)

ggplot(train, aes(x = Discount, y = margin)) +
  geom_point(color="#80b1d3", size = 3, alpha=.9) +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) +
  labs(x = "Discount", y = "Avg. Gross Margin") + theme_minimal()+
  geom_abline(slope = coef(lm.fit)[["Discount"]], 
              intercept = coef(lm.fit)[["(Intercept)"]])
```

```{r, message = FALSE}
pred <- predict(lm.fit, newdata = test)
rmse <- sqrt(sum((pred - test$margin)^2)/length(test$margin))
print(paste("RMSE =", rmse))
normalized_rmse <- 100 * rmse/(max(train$margin)-min(train$margin))
cat(paste("RMSE =", rmse))
cat(paste("\nNormalized RMSE =", normalized_rmse, "%"))
```

Diagnostic plots

```{r, message = FALSE, fig.height=10, fig.width=10}

par(mfrow = c(2, 2))
plot(lm.fit)
```

```{r, message = FALSE, fig.height=10, fig.width=10}
par(mfrow = c(2, 2))
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

## Multiple linear regression

Create training and test dataset

```{r, message = FALSE}
set.seed(1)
row.number <- sample(1:nrow(df), 0.8*nrow(df))
train = df[row.number,]
test = df[-row.number,]
dim(train)
dim(test)
```

```{r, message = FALSE}
lm_mult = train(
  form = Profit ~ Sales + Quantity + Segment + Discount + Processing.Time + Ship.Mode + Region + Sub.Category + Gross.Margin,
  data = train,
  trControl = trainControl(method = "cv", number = 5),
  method = "lm"
)

lm_mult
lm_mult$results
summary(lm_mult)
```

Check VIF:

```{r, message = FALSE}
vif(lm_mult$finalModel) 
```

Multicolinearity in Discount and Gross Margin. Remove Gross Margin

Plots:

```{r, message = FALSE, fig.height=10, fig.width=10}
model = lm(Profit ~ Sales + Quantity + Segment + Discount + Processing.Time + Ship.Mode + Region + Sub.Category, data = train)
par(mfrow=c(2,2))
plot(model)
```

### Interaction terms

```{r, message = FALSE}
# Interactions to check:
# Sales:Quantity -----> No improvement in the model
# Sales:Discount -------------> Increases a lot the adjusted R squared, good. Vif also okay
# Ship.Mode:Processing.Time --------------- No improvement
# Sales:Sub.Category -------- Adjusted R squared increases but vif factors are not good.
# Sales:Segment -----------------------> Slight increase in R^2 and vif factors, but still okay. Won't take, since doesn't improve R so much.
# Region:Sales ---->   Increases a bit the adjusted R squared but also vif factors, so better not to take.

# So we keep the interaction between Sales and Discount and substitute Discount
# by Gross.Margin
lm_mult.fit <- lm(Profit ~ Sales + Quantity + Segment + Gross.Margin + Processing.Time + Sales:Discount + Ship.Mode + Region + Sub.Category , data = train)
summary(lm_mult.fit)
vif(lm_mult.fit)
```

Use anova to study Sales:Discount

The null hypothesis is that the two models fit the data equally well, and the alternative hypothesis is that the full model is superior

```{r, message = FALSE}
lm_mult.fit2 <- lm(Profit ~ Sales + Sales:Discount + Quantity + Segment + Gross.Margin + Processing.Time + Ship.Mode + Region + Sub.Category , data = train)
lm_mult.fit <- lm(Profit ~ Sales + Quantity + Segment + Discount + Processing.Time + Ship.Mode + Region + Sub.Category , data = train)
anova(lm_mult.fit , lm_mult.fit2)
```

### Nonlinear terms

```{r, message = FALSE}
# Quantity^2 -----> Good pvalue but high vif factor and no improvement of R
# Discount^2 ------------> high pvalue
# Poly(Sales, 3) ------------>  good pvalue and increase in R, no problem with vif
# Poly(Quantity,5) ------------> Nope, only Quantity^2 p-value is good, and we have seen before vif are high
# log(sales) ------------> adding this term to Sales has a good p value, but no improvement on R. Replacing the Sales by log(Sales) very bad R^2
# sqrt(Sales) ------------> same as with log(sales)
# log(Quantity) ------------> same problem with log(sales) but anova shows is not that good
# sqrt(Quantity) ------------> same as log(quantity) but with pvalue in anova 3%
#sqrt(Gross.Margin) ------------> good p-value, good increase in adjusted R, good vif

lm_mult.fit <- lm(Profit ~ poly(Sales, 3) + Sales:Discount + Segment + Gross.Margin + Processing.Time + Ship.Mode + Region + Sub.Category , data = train)
summary(lm_mult.fit)
vif(lm_mult.fit)
```

Use anova to study Poly(Sales, 3) The null hypothesis is that the two models fit the data equally well, and the alternative hypothesis is that the full model is superior

```{r, message = FALSE}
lm_mult.fit2 <- lm(Profit ~ poly(Sales, 3) + Quantity + Segment + Gross.Margin + Processing.Time + Ship.Mode + Region + Sub.Category , data = train)
lm_mult.fit <- lm(Profit ~ Sales + Quantity + Segment + Gross.Margin + Processing.Time + Ship.Mode + Region + Sub.Category , data = train)

anova(lm_mult.fit , lm_mult.fit2)
```

### Lasso regression

```{r, message = FALSE}
profit <- train$Profit
predictors <- data.matrix(train[, c("Sales", "Quantity", "Segment", "Discount", "Processing.Time", "Ship.Mode", "Region", "Sub.Category", "Gross.Margin")])
```

Find best lambda

```{r, message = FALSE, fig.height=5, fig.width=5}
#perform k-fold cross-validation to find optimal lambda value (alpha = 0 is ridge, alpha = 1 is lasso)
cv_model <- cv.glmnet(predictors, profit, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda

#produce plot of test MSE by lambda value
plot(cv_model) 
```

Find coefficients of best model

```{r, message = FALSE}
best_model <- glmnet(predictors, profit, alpha = 1, lambda = best_lambda)
coef(best_model)
```

### Ridge regression

```{r, message = FALSE}
profit <- train$Profit
predictors <- data.matrix(train[, c("Sales", "Quantity", "Segment", "Discount", "Processing.Time", "Ship.Mode", "Region", "Sub.Category", "Gross.Margin")])
```

Find best lambda

```{r, message = FALSE, fig.height=5, fig.width=5}
#perform k-fold cross-validation to find optimal lambda value (alpha = 0 is ridge, alpha = 1 is lasso)
cv_model <- cv.glmnet(predictors, profit, alpha = 0)

#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda

#produce plot of test MSE by lambda value
plot(cv_model) 
```

Find coefficients of best model

```{r, message = FALSE}
best_model <- glmnet(predictors, profit, alpha = 0, lambda = best_lambda)
coef(best_model)
```

### Final variable selection

Based on previous analysis of Lasso and Ridge regression, interaction terms and nonlinear terms let's create a new model selecting the most important variables Give final result with 5-fold cv

```{r, message = FALSE, fig.height=10, fig.width=10}
lm_mult = train(
  form = Profit ~ poly(Sales,3) + Sales:Discount + Sub.Category + Gross.Margin, # Also deleted Region because found to have high p-value in the new model.
  data = train,
  trControl = trainControl(method = "cv", number = 5),
  method = "lm")


lm_mult
lm_mult$results
summary(lm_mult)
```

Check VIF:

```{r, message = FALSE, fig.height=10, fig.width=10}
vif(lm_mult$finalModel)
# Plot:
par(mfrow=c(2,2))
plot(lm_mult$finalModel)
```

```{r, message = FALSE}
pred <- predict(lm_mult, newdata = test)
rmse <- sqrt(sum((pred - test$Profit)^2)/length(test$Profit))
normalized_rmse <- 100 * rmse/(max(train$Profit)-min(train$Profit))
cat(paste("RMSE =", rmse))
cat(paste("\nNormalized RMSE =", normalized_rmse, "%"))
```

```{r, message = FALSE, fig.height=5, fig.width=5}
plot(test$Profit, pred)
```

## Classification

In this section we will address a binary classification problem. The idea is to make the column *Profit* a binary column, so that all positive profits get label *Positive* and all negative profits get label *Negative*. By doing this we can train a model that is able to predict if a purchase will result in positive or negative profit based on features such as Region, State, City, Sub-Category, Segment, Ship.Mode, Sales, Quantity, Discount and Process time.

As a first step let us define a function to compute the most typical evaluation metrics: accuracy, precision, recall and F1 score.

```{r, message = FALSE}
calc_metrics <- function(cm) {

  n = sum(cm) # number of instances
  nc = nrow(cm) # number of classes
  diag = diag(cm) # number of correctly classified instances per class 
  rowsums = rowSums(cm) # number of instances per class
  colsums = colSums(cm) # number of predictions per class
  p = rowsums / n # distribution of instances over the actual classes
  q = colsums / n # distribution of instances over the predicted classes

  accuracy = sum(diag) / n
  precision = diag / colsums 
  recall = diag / rowsums 
  f_1 = 2 * precision * recall / (precision + recall) 
   
  print(paste("Accuracy", accuracy))
  print("Precision: ")
  print(precision)
  print("Recall: ")
  print(recall)
  print("F1 score: ")
  print(f_1)

}
```

The column *Profit* is then modified and the discrete value features are converted to factors.

```{r, message = FALSE}
df_log <- df[, c("Profit","Discount", "Quantity", "Sales", "Segment", "Ship.Mode", "Region", "Order.Month", "Processing.Time", "Sub.Category", "Order.Year")]
df_log["Profit"]<-replace(df_log["Profit"], df["Profit"] < 0, "Negative")
df_log["Profit"]<-replace(df_log["Profit"], df["Profit"] >= 0, "Positive")
df_log[,"Profit"]<- as.factor(df_log[,"Profit"])
df_log[,"Segment"]<- as.factor(df_log[,"Segment"])
df_log[,"Ship.Mode"]<- as.factor(df_log[,"Ship.Mode"])
df_log[,"Region"]<- as.factor(df_log[,"Region"])
df_log[,"Sub.Category"]<- as.factor(df_log[,"Sub.Category"])
```

Let's check for unbalance in the data.

```{r, message = FALSE}
table(df_log$Profit)
prop.table(table(df_log$Profit))
contrasts(df_log$Profit)
```

As we can see there is moderate unbalance in our data. We are going to tackle this problem by undersampling and oversampling accordingly. But first let's see what results we get with the unbalanced data. The train-test split is done taking as training all examples corresponding to the years 2014, 2015 and 2016, leaving 2017 as the test set.

```{r, message = FALSE}
train <- (df_log$Order.Year < 2017)
df_trn <- df_log[train, ]
df_tst <- df_log[!train, ]
dim(df_tst)
Profit.2017 <- df_log$Profit[!train]
```

### Feature selection

Before training the model feature selection is done to avoid overfitting and thus ensure better results. The methodology followed to select the best features is based on the comparison of the AIC and the McFadden's $R^{2}$ for different trained linear models obtained by suppressing features one at a time.

The Akaike Information Criterion (AIC) is a mathematical method for evaluating how well a model fits the data it was generated from.

Since GLM models use a maximum likelihood estimator, there is no minimization of the squared error and hence no *R* is calculated. There are a large number of pseudo-$R^{2}$ for GLMs, among which the most popular is McFadden's $R^2$. It is calculated as:

$$R^{2} = 1 - \frac{Residual Deviance}{Null Deviance}$$ We will therefore look for models that present a bigger McFadden's $R^{2}$.

```{r, message = FALSE}
# Let's try now doing the same but with less variables (The ones with higher p value in the full fitted model)

# glm.fits1 <- glm(Profit ~  Discount + Quantity + Sales + Ship.Mode + Sub.Category , data = df_trn, family = binomial)
# glm.fits2 <- glm(Profit ~  Discount + Quantity + Sales + Ship.Mode +  Processing.Time , data = df_trn, family = binomial)
# glm.fits3 <- glm(Profit ~  Discount + Sales +  Sub.Category , data = df_trn, family = binomial)
# glm.fits4 <- glm(Profit ~  Discount + Quantity + Sales + Ship.Mode + Region + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits5 <- glm(Profit ~  Discount + Sales  + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits6 <- glm(Profit ~  Discount + Quantity + Sales + Region + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits7 <- glm(Profit ~  Discount + Quantity + Sales + Ship.Mode + Region + Order.Month + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits8 <- glm(Profit ~  Discount +  Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits9 <- glm(Profit ~  Discount + Sales + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits10 <- glm(Profit ~ Quantity + Sales + Ship.Mode + Region + Order.Month + Processing.Time + Sub.Category , data = df_trn, family = binomial)


glm.fits1 <- glm(Profit ~  Discount + Quantity + Sales + Segment + Ship.Mode + Region + Order.Month + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits2 <- glm(Profit ~  Discount +  Sales + Ship.Mode + Region + Processing.Time , data = df_trn, family = binomial) # Subcategory is important
glm.fits3 <- glm(Profit ~  Discount + Sales + Ship.Mode + Region +  Sub.Category , data = df_trn, family = binomial)
glm.fits4 <- glm(Profit ~  Discount + Sales +  Ship.Mode + Region  + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits5 <- glm(Profit ~  Discount +  Sales +  Ship.Mode + Processing.Time + Sub.Category , data = df_trn, family = binomial) # Region seems to be important
# glm.fits6 <- glm(Profit ~  Discount + Sales +  Region + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits7 <- glm(Profit ~  Discount +  Sales + Ship.Mode + Region + Processing.Time + Sub.Category , data = df_trn, family = binomial)
glm.fits8 <- glm(Profit ~  Discount +  Ship.Mode + Region + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits9 <- glm(Profit ~  Discount + Sales +  Ship.Mode + Region + Processing.Time + Sub.Category , data = df_trn, family = binomial)
# glm.fits10 <- glm(Profit ~  Quantity + Sales +  Ship.Mode + Region + Order.Month + Processing.Time + Sub.Category , data = df_trn, family = binomial) # Discount is very important
```

```{r, message = FALSE}
print(paste("AIC1: ", extractAIC(glm.fits1)[2]))
print(paste("McFadden's R^2  ", with(summary(glm.fits1), 1 - deviance/null.deviance)))

# print(paste("AIC2: ", extractAIC(glm.fits2)[2]))
# print(paste("McFadden's R^2  ", with(summary(glm.fits2), 1 - deviance/null.deviance)))

print(paste("AIC3: ", extractAIC(glm.fits3)[2]))
print(paste("McFadden's R^2  ", with(summary(glm.fits3), 1 - deviance/null.deviance)))

print(paste("AIC4: ", extractAIC(glm.fits4)[2]))
print(paste("McFadden's R^2  ", with(summary(glm.fits4), 1 - deviance/null.deviance)))

# print(paste("AIC5: ", extractAIC(glm.fits5)[2]))
# print(paste("McFadden's R^2  ", with(summary(glm.fits5), 1 - deviance/null.deviance)))

# print(paste("AIC6: ", extractAIC(glm.fits6)[2]))
# print(paste("McFadden's R^2  ", with(summary(glm.fits6), 1 - deviance/null.deviance)))

# print(paste("AIC7: ", extractAIC(glm.fits7)[2]))
# print(paste("McFadden's R^2  ", with(summary(glm.fits7), 1 - deviance/null.deviance)))

print(paste("AIC8: ", extractAIC(glm.fits8)[2]))
print(paste("McFadden's R^2  ", with(summary(glm.fits8), 1 - deviance/null.deviance)))

# print(paste("AIC9: ", extractAIC(glm.fits9)[2]))
# print(paste("McFadden's R^2  ", with(summary(glm.fits9), 1 - deviance/null.deviance)))

# print(paste("AIC10: ", extractAIC(glm.fits10)[2]))
# print(paste("McFadden's R^2  ", with(summary(glm.fits10), 1 - deviance/null.deviance)))
```

Best features found are: Discount, Sales, Sub.Category, Processing.Time and Region. We choose only those 5 to keep the problem simple and because the addition of other features does not improve the performance of the model.

### Logistic regression

#### Unbalanced Dataset and Crossvalidation

```{r, message = FALSE}
glm_mod = train(
  form = Profit ~ Discount + Sales + Sub.Category + Region + Processing.Time,
  data = df_trn,
  trControl = trainControl(method = "cv", number = 5),
  method = "glm",
  family = "binomial"
)

glm_mod
glm_mod$results
```

```{r, message = FALSE}
summary(glm_mod)
print("McFadden's R^2")
with(summary(glm_mod), 1 - deviance/null.deviance)
```

```{r, message = FALSE}
glm.probs <- predict(glm_mod, df_tst)
levels = c("Positive", "Negative")
cm <- as.matrix(table(glm.probs , Profit.2017)[levels,levels])

cm
calc_metrics(cm)
roc.curve(Profit.2017, glm.probs, plotit = F)
```

#### Balanced Dataset and Crossvalidation

```{r, message = FALSE}
balanced_data <- ROSE(Profit ~ ., data = df_trn, seed = 1)$data
table(balanced_data$Profit)
contrasts(balanced_data$Profit)
```

```{r, message = FALSE}
glm_fits = train(
  form = Profit ~ Discount + Sales + Sub.Category + Region + Processing.Time,
  data = balanced_data,
  trControl = trainControl(method = "cv", number = 5),
  method = "glm",
  family = "binomial"
)

glm_fits
glm_fits$results
```

```{r, message = FALSE}
summary(glm_fits)
print("McFadden's R^2")
with(summary(glm_fits), 1 - deviance/null.deviance)
```

```{r, message = FALSE}
glm.probs <- predict(glm_fits , df_tst)
levels = c("Positive", "Negative")
cm <- as.matrix(table(glm.probs, Profit.2017)[levels,levels])

cm
calc_metrics(cm)
roc.curve(Profit.2017, glm.probs, plotit = F)
```

Let's try to balance the data with another method

```{r, message = FALSE}
balanced_data <- ovun.sample(Profit ~ ., data = df_trn, method = "both", p=0.5, N = nrow(df_trn), seed = 1)$data # Can play with parameter p: probability of positive class in newly generated sample
table(balanced_data$Profit)
```

```{r, message = FALSE}
glm_fits = train(
  form = Profit ~ Discount + Sales + Sub.Category + Region + Processing.Time,
  data = balanced_data,
  trControl = trainControl(method = "cv", number = 5),
  method = "glm",
  family = "binomial"
)

glm_fits
glm_fits$results
```

```{r, message = FALSE}
summary(glm_fits)
print("McFadden's R^2")
with(summary(glm_fits), 1 - deviance/null.deviance)
```

```{r, message = FALSE}
glm.probs <- predict(glm_fits , df_tst)
levels = c("Positive", "Negative")
cm <- as.matrix(table(glm.probs, Profit.2017)[levels,levels])

cm
calc_metrics(cm)
roc.curve(Profit.2017, glm.probs, plotit = F)
```

### Bayes classifier with Crossvalidation

```{r, message = FALSE}
lda_mod = train(
  form = Profit ~ Discount + Sales + Region + Processing.Time + Sub.Category,
  data = df_trn,
  trControl = trainControl(method = "cv", number = 5),
  method = "lda"
)

lda_mod
lda_mod$results
```

```{r, message = FALSE}
lda.probs <- predict(lda_mod, df_tst)
levels = c("Positive", "Negative")
cm <- as.matrix(table(lda.probs , Profit.2017)[levels,levels])
cm
calc_metrics(cm)
roc.curve(Profit.2017, lda.probs, plotit = F)
```

### Linear discriminant analysis

#### Unbalanced Dataset and Crossvalidation

```{r, message = FALSE}
lda_mod = train(
  form = Profit ~ Discount + Sales + Region + Processing.Time + Sub.Category,
  data = df_trn,
  trControl = trainControl(method = "cv", number = 5),
  method = "lda"
)

lda_mod
lda_mod$results
```

```{r, message = FALSE}
lda.probs <- predict(lda_mod, df_tst)
levels = c("Positive", "Negative")
cm <- as.matrix(table(lda.probs , Profit.2017)[levels,levels])
cm
calc_metrics(cm)
roc.curve(Profit.2017, lda.probs, plotit = F)
```

#### Balanced Dataset and Crossvalidation

```{r, message = FALSE}
lda_mod = train(
  form = Profit ~ Discount + Sales + Region + Processing.Time + Sub.Category,
  data = balanced_data,
  trControl = trainControl(method = "cv", number = 5),
  method = "lda"
)

lda_mod
lda_mod$results
```

```{r, message = FALSE}
lda.probs <- predict(lda_mod, df_tst)
levels = c("Positive", "Negative")
cm <- as.matrix(table(lda.probs , Profit.2017)[levels,levels])
cm
calc_metrics(cm)
roc.curve(Profit.2017, lda.probs, plotit = F)
```

### Quadratic discriminant analysis

```{r, message = FALSE}
qda_mod = train(
  form = Profit ~ Discount + Sales + Region + Processing.Time,
  data = df_trn,
  trControl = trainControl(method = "cv", number = 5),
  method = "qda"
)

qda_mod
qda_mod$results
```

```{r, message = FALSE}
qda.probs <- predict(qda_mod, df_tst)
levels = c("Positive", "Negative")
cm <- as.matrix(table(qda.probs , Profit.2017)[levels,levels])
cm
calc_metrics(cm)
roc.curve(Profit.2017, qda.probs, plotit = F)
```
